---
title: "Dataset Characteristics and Diagnostics"
output: html_document
date: "`r Sys.Date()`"
params:
  species: "Vehkaoja_Dog"
  wd: "C:/Users/oaw001/OneDrive - University of the Sunshine Coast/DataReformatting"
  sample_rate: "100"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# strip the params
#setwd(params$wd)
#sample_rate <- as.numeric(params$sample_rate)
species <- "Annett_Glider"
sample_rate <- 20

# packages
pacman::p_load(
  data.table,
  tidyverse,
  patchwork
  )

# plotting stuff
my_theme <- function() {
  theme_minimal(base_size = 12) +
    theme(
      panel.border = element_rect(color = "black", linewidth = 1.5, fill = NA),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      plot.background = element_blank(),
      panel.background = element_blank(),
      axis.line = element_line(color = "black"),
      axis.ticks = element_line(color = "black")
    )
}

my_colours = c("coral", "aquamarine3", "orchid3", "slateblue2", "goldenrod2", "deepskyblue3", "firebrick3", "tan1", "lightcoral" , "darkcyan", "tomato", "plum")

# load in the data
data <- fread(file.path("C:/Users/norri/OneDrive/Documents/GitHub/DataReformatting", species, paste0(species, "_formatted.csv")))


```

This a report showing basic at-a-glance diagnostics for each dataset.

## Metadata
```{r}
# Table showing the metadata
# TODO: please make this :)
```

## Data volume
How much data is there, per activity and per ID.

```{r volume of data, echo=FALSE,  fig.width = 10, fig.height = 4}

vol <- data %>% 
  group_by(ID, Activity) %>%
  count() %>%
  mutate(mins = (n/sample_rate)/60)

act_vol_plot <- ggplot(vol, aes(x = Activity, y = mins, fill = as.factor(ID))) +
  geom_col() +
  labs(x = "Activity", y = "Volume of data (mins)") +
  my_theme() + theme(legend.position = "none", 
                     axis.text.x = element_text(angle = 90, hjust = 1, vjust = 1))

ID_vol_plot <- ggplot(vol, aes(x = as.factor(ID), y = mins, fill = as.factor(Activity))) +
  geom_col() +
  labs(x = "ID", y = "Volume of data (mins)") +
  my_theme() + theme(legend.position = "none",
                     axis.text.x = element_text(angle = 90, hjust = 1, vjust = 1))

act_vol_plot + ID_vol_plot

```

## Sequences
How much of the data is collected in sequence, how many transitions are captured. Plot 1 -> length of sequences. Plot 2 -> number of transtions in the sequence

```{r sequences, echo = FALSE,  fig.width = 10, fig.height = 4}
seq_len <- data %>%
  group_by(ID, sequence) %>%
  count() %>%
  mutate(mins = (n/sample_rate)/60)

seq_len_plot <- ggplot(seq_len, aes(x = mins, fill = as.factor(ID))) +
 geom_histogram() + 
  labs(x = "Length of continuous sequence (min)", y = "Frequency") +
  my_theme() + theme(legend.position = "none") 

# now looking at the number of transitions
data <- data %>%
  group_by(ID, sequence) %>%
  arrange(Time) %>%
  mutate(
    change_point = if_else(lag(Activity) == Activity, 0L, 1L),
    change_point = replace_na(change_point, 0L),
    event = cumsum(change_point)
  ) %>%
  select(-change_point) %>%
  ungroup() %>%
  arrange(ID, Time)

seq_trans <- data %>%
  group_by(ID, sequence) %>%
  summarise(transitons = max(event))

seq_trans_plot <- ggplot(seq_trans, aes(x = transitons, fill = as.factor(ID))) +
 geom_histogram() + 
  labs(x = "Number of behavioural transitions captured within the sequence", y = "Frequency") +
  my_theme() + theme(legend.position = "none")

seq_len_plot + seq_trans_plot

```

## Durations
What was the duration of the behaviours

```{r durations, echo=FALSE, fig.width = 10, fig.height = 4}

train_lengths <- data %>%
  group_by(ID, sequence, Activity, event) %>%
  count() %>%
  mutate(mins = (n/sample_rate)/60)

lengths <- ggplot(train_lengths, aes(x = as.factor(Activity), y = mins)) +
  geom_boxplot() +
  labs(x = "Activity", y = "Length of each individual event (mins)") +
  my_theme() + theme(legend.position = "none",
                     axis.text.x = element_text(angle = 90, hjust = 1, vjust = 1))

# make a distribution of the behaviours
freqs <- ggplot(train_lengths, aes(x = mins, fill = as.factor(event))) +
  geom_histogram() + 
  labs(x = "Length of each individual event (mins)", y = "Frequency") +
  my_theme() +  theme(legend.position = "none")

lengths + freqs

# and now actually just print the stats
stat_mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

# its in raw samples at the moment so need to summarise to seconds
train_summary <- train_lengths %>%
  group_by(Activity) %>%
  summarise(
    min_len= min(mins),
    mean   = mean(mins),
    median = median(mins),
    mode   = stat_mode(mins),
    p75    = quantile(mins, 0.75),
    p95    = quantile(mins, 0.95),
    max    = max(mins),
    .groups = "drop"
  )

# TODO: is there a better way to display this information??
print(train_summary)

```

## Behavioural Samples
Visualising what each of the behaviours look like.

```{r trace examples, echo = FALSE, fig.width = 10, fig.height = 10}
plotTraceExamples <- function(behaviours, data, individuals, n_samples) {
  
  data <- data %>% filter(ID %in% sample(unique(data$ID), individuals))
  
  # Create plots for each behavior (with error catching)
  plots <- purrr::map(behaviours, function(behaviour) {
    tryCatch(
      {
        plot_behaviour(behaviour, n_samples, data)
      },
      error = function(e) {
        message("Skipping plot for ", behaviour, ": ", e$message)
        NULL  # Return NULL to indicate skipping
      }
    )
  })
  
  # Remove NULL plots (for behaviors with no data)
  plots <- purrr::compact(plots)
  
  # Combine plots into a single grid
  grid_plot <- cowplot::plot_grid(plotlist = plots)
  
  return(list(plots = plots, 
                 grid_plot = grid_plot))
}

# Function to create the plot for each behavior
plot_behaviour <- function(behaviour, n_samples, data) {
  df <- data %>%
    filter(Activity == behaviour) %>%
    group_by(ID, Activity) %>%
    slice(1:n_samples) %>%
    mutate(relative_time = row_number())
  
  # Check if the filtered dataframe is empty
  if (nrow(df) == 0) {
    stop("No data available for behaviour: ", behaviour)
  }
  
  ggplot(df, aes(x = relative_time)) +
    geom_line(aes(y = X, color = "X"), show.legend = FALSE) +
    geom_line(aes(y = Y, color = "Y"), show.legend = FALSE) +
    geom_line(aes(y = Z, color = "Z"), show.legend = FALSE) +
    labs(title = paste(behaviour),
         x = NULL, y = NULL) +
    scale_color_manual(values = c(X = "aquamarine3", Y = "plum", Z = "goldenrod"), guide = "none") +
    facet_wrap(~ ID, nrow = 1, scales = "free_x") +
    theme_minimal() +
    theme(panel.grid = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank())
}


# data <- fread(file.path("C:/Users/oaw001/OneDrive - University of the Sunshine Coast/DataReformatting/Smit_Cat/Smit_Cat_Labelled.csv")) %>%
#   rename(X = Accelerometer.X,
#          Y = Accelerometer.Y,
#          Z = Accelerometer.Z,
#          Time = datetime_precise)
# cahnge these variables to change what is plotted
tracePlots <- plotTraceExamples(
  behaviours = unique(data$Activity),
  data = data,
  individuals = ifelse(length(unique(data$ID))>5, 5, length(unique(data$ID))),
  if(sample_rate <5) {
    n_samples = 6 * sample_rate
  } else {
    n_samples = 2 * sample_rate
  }
)

 



tracePlots$grid_plot

```

